\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}

### 2014-12-03 papers/jss/article.Rnw
Improve the plot and point out 3 outliers now and explain them in the text.  Correct an error in the space savings definition.  Change trivial example to simple example.
Suggestions from: Andy Chu


git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@927 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Figure~\ref{fig:compression} shows the space savings $\left(1 - \frac{\textrm{Uncompressed Size}}{\textrm{Compressed Size}}\right)$ for each of the data sets using each of these four methods.  The associated table shows the exact data sizes for two outliers and the aggregate of all \Sexpr{n} data sets.
Note that Protocol Buffer serialization results in slightly
smaller byte streams compared to native \proglang{R} serialization in most cases,
but this difference disappears if the results are compressed with gzip.
%Sizes are comparable but Protocol Buffers provide simple getters and setters
%in multiple languages instead of requiring other programs to parse the \proglang{R}
%serialization format. % \citep{serialization}.

The \code{crimtab} dataset of anthropometry measurements of British
prisoners \citep{garson1900metric}
shows the greatest difference in the space savings when
using Protocol Buffers compared to \proglang{R} native serialization.
This dataset is a 42x22 table of integers, most equal to 0.  Small
integer values like this can be very efficiently encoded by the
\emph{Varint} integer encoding scheme used by Protocol Buffers which
use a variable number of bytes for each value.

@@

\begin{figure}[t!]
\begin{center}
<<echo=FALSE,fig=TRUE,width=8,height=4>>=
plot(clean.df$savings.serialized, clean.df$savings.rprotobuf, pch=1, col="red", las=1, xlab="Serialization Space Savings", ylab="Protocol Buffer Space Savings")
points(clean.df$savings.serialized.gz, clean.df$savings.rprotobuf.gz,pch=2, col="blue")
# grey dotted diagonal
abline(a=0,b=1, col="grey",lty=2,lwd=3)

@@

# The one to label.
tmp.df <- clean.df[which(clean.df$savings.diff == min(clean.df$savings.diff)),]
# This minimum means most to the left of our line, so pos=2 is label to the left
text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=2)
text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=2)

tmp.df <- clean.df[which(clean.df$savings.diff == max(clean.df$savings.diff)),]
# This minimum means most to the right of the diagonal, so pos=4 is label to the right
text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=4)
text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=4)

@@

legend("topleft", c("Raw", "Gzip Compressed"), pch=1:2, col=c("red", "blue"))

interesting.df <- clean.df[unique(c(which(clean.df$savings.diff == min(clean.df$savings.diff)),
                             which(clean.df$savings.diff == max(clean.df$savings.diff)),
                             which(clean.df$savings.diff.gz == max(clean.df$savings.diff.gz)),
			     which(clean.df$dataset == "TOTAL"))),c("dataset", "object.size", "serialized", "gzipped serialized", "RProtoBuf", "gzipped RProtoBuf", "savings.serialized", "savings.serialized.gz", "savings.rprotobuf", "savings.rprotobuf.gz")]
# Print without .00 in xtable
interesting.df$object.size <- as.integer(interesting.df$object.size)
@

% latex table generated in R 3.0.2 by xtable 1.7-0 package
% Wed Nov 26 15:31:30 2014
%\begin{table}[ht]
%\begin{center}
\begin{tabular}{rlrrrrr}
  \toprule
  Data Set & object.size & \multicolumn{2}{c}{\proglang{R} Serialization} &
  \multicolumn{2}{c}{RProtoBuf Serialization} \\
  & & default & gzipped & default & gzipped \\
  \cmidrule(r){2-6}
 crimtab & 7,936 & 4,641 (41.5\%) & 713 (91.0\%) & 1,655 (79.2\%) & 576 (92.7\%)\\
 faithful & 5,136 & 4,543 (11.5\%) & 1,339 (73.9\%) & 4,936 (3.9\%) & 1,776 (65.5\%)\\
   \hline
 All & 605,256 & 461,667 (24\%) & 138,937 (77\%) & 435,360 (28\%) & 142,134 (77\%)\\
\hline
\end{tabular}
\end{center}
\caption{(Top) Relative space savings of Protocol Buffers and native \proglang{R} serialization over the raw object sizes of each of the \Sexpr{n} data sets in the \pkg{datasets} package. Points to the left of the dashed $y=x$ line represent datasets that are more efficiently encoded with Protocol Buffers. (Bottom) Absolute space savings of two outlier datasets and the aggregate performance of all datasets.}
\label{fig:compression}
\end{figure}

@@

\subsection*{A trivial single-machine example for Python to R serialization}

>>>>

Figure~\ref{fig:compression} shows the space savings $\left(1 - \frac{\textrm{Compressed Size}}{\textrm{Uncompressed Size}}\right)$ for each of the data sets using each of these four methods.  The associated table shows the exact data sizes for some outliers and the aggregate of all \Sexpr{n} data sets.
Note that Protocol Buffer serialization results in slightly
smaller byte streams compared to native \proglang{R} serialization in most cases (red dots),
but this difference disappears if the results are compressed with gzip (blue triangles).
%Sizes are comparable but Protocol Buffers provide simple getters and setters
%in multiple languages instead of requiring other programs to parse the \proglang{R}
%serialization format. % \citep{serialization}.

The \code{crimtab} dataset of anthropometry measurements of British
prisoners \citep{garson1900metric} and the \code{airquality} dataset of air quality measurements in New York show the
greatest difference in the space savings when
using Protocol Buffers compared to \proglang{R} native serialization.
The \code{crimtab} dataset is a 42x22 table of integers, most equal to 0, and the \code{airquality} dataset is a data frame of 154 observations of 1 numeric and 5 integer variables.  In both data sets, the large number of small
integer values can be very efficiently encoded by the
\emph{Varint} integer encoding scheme used by Protocol Buffers which
use a variable number of bytes for each value.

@@

\begin{figure}[hbt!]
\begin{center}
<<label=SER,echo=FALSE,include=FALSE,fig=TRUE>>=
old.mar<-par("mar")
new.mar<-old.mar
new.mar[3]<-0
new.mar[4]<-0
my.cex<-1.3
par("mar"=new.mar)
plot(clean.df$savings.serialized, clean.df$savings.rprotobuf, pch=1, col="red", las=1, xlab="Serialization Space Savings", ylab="Protocol Buffer Space Savings", xlim=c(0,1),ylim=c(0,1),cex.lab=my.cex, cex.axis=my.cex)
points(clean.df$savings.serialized.gz, clean.df$savings.rprotobuf.gz,pch=2, col="blue")
# grey dotted diagonal
abline(a=0,b=1, col="grey",lty=2,lwd=3)

@@

# The one to label.
tmp.df <- clean.df[which(clean.df$savings.diff == min(clean.df$savings.diff)),]
# This minimum means most to the left of our line, so pos=2 is label to the left
text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=2, cex=my.cex)

# Some gziped version
# text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=2, cex=my.cex)

# Second one is also an outlier
tmp.df <- clean.df[which(clean.df$savings.diff == sort(clean.df$savings.diff)[2]),]
# This minimum means most to the left of our line, so pos=2 is label to the left
text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=2, cex=my.cex)
#text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=my.cex)

tmp.df <- clean.df[which(clean.df$savings.diff == max(clean.df$savings.diff)),]
# This minimum means most to the right of the diagonal, so pos=4 is label to the right
# Only show the gziped one.
#text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=4, cex=my.cex)
text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=4, cex=my.cex)

@@

legend("topleft", c("Raw", "Gzip Compressed"), pch=1:2, col=c("red", "blue"), cex=my.cex)

interesting.df <- clean.df[unique(c(which(clean.df$savings.diff == min(clean.df$savings.diff)),
                             which(clean.df$savings.diff == max(clean.df$savings.diff)),
                             which(clean.df$savings.diff.gz == max(clean.df$savings.diff.gz)),
			     which(clean.df$dataset == "TOTAL"))),c("dataset", "object.size", "serialized", "gzipped serialized", "RProtoBuf", "gzipped RProtoBuf", "savings.serialized", "savings.serialized.gz", "savings.rprotobuf", "savings.rprotobuf.gz")]
# Print without .00 in xtable
interesting.df$object.size <- as.integer(interesting.df$object.size)
par("mar"=old.mar)
@
\includegraphics[width=0.45\textwidth]{figures/fig-SER}

% latex table generated in R 3.0.2 by xtable 1.7-0 package
% Wed Nov 26 15:31:30 2014
%\begin{table}[ht]
%\begin{center}
\begin{tabular}{rlrrrrr}
  \toprule
  Data Set & object.size & \multicolumn{2}{c}{\proglang{R} Serialization} &
  \multicolumn{2}{c}{RProtoBuf Serialization} \\
  & & default & gzipped & default & gzipped \\
  \cmidrule(r){2-6}
 crimtab & 7,936 & 4,641 (41.5\%) & 713 (91.0\%) & 1,655 (79.2\%) & 576 (92.7\%)\\
 airquality & 5,496 & 4,551 (17.2\%) & 1,241 (77.4\%) & 2,874 (47.7\%) & 1,294 (76.5\%)\\
 faithful & 5,136 & 4,543 (11.5\%) & 1,339 (73.9\%) & 4,936 (3.9\%) & 1,776 (65.5\%)\\
   \hline
 All & 605,256 & 461,667 (24\%) & 138,937 (77\%) & 435,360 (28\%) & 142,134 (77\%)\\
\hline
\end{tabular}
\end{center}
\caption{(Top) Relative space savings of Protocol Buffers and native \proglang{R} serialization over the raw object sizes of each of the \Sexpr{n} data sets in the \pkg{datasets} package. Points to the left of the dashed $y=x$ line represent datasets that are more efficiently encoded with Protocol Buffers. (Bottom) Absolute space savings of three outlier datasets and the aggregate performance of all datasets.}
\label{fig:compression}
\end{figure}

@@

\subsection*{A simple single-machine example for Python to R serialization}
